---
title: "Pulling ACS Data for Inspace Projects"
author: "Amy Youngbloom"
date: "11/2/2022"
output: html_document
---

```{r setup, include=FALSE}
#knit this file with the working directory of "workspace/" instead of "workspace/examples"
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, eval=FALSE)
knitr::opts_knit$set(root.dir = '/home/rstudio/workspace/')
```

### Introduction

To allow for pooling of Inspace projects, we have generated a wishlist of ACS variables that we are hoping each project can pull for all participants. ACS variables will be pulled for 2017 (2013-2017 5-year ACS) as well as for the 5-year period centered around your initial year of enrollment. For example, if enrollment in your program began in 2016, you would want to pull data from the 2014-2019 ACS (year=2019). 

This document lays out instructions for how to pull the requested ACS variables for each participant. 

```{r load packages & ACMT}
library(tidycensus)
library(tidyverse)
library(dplyr)
#install.packages('janitor')
library(janitor)
library(reshape2)
source("GeocoderACMT.R")
source('Inspace/Inspace_external_data_functions.R')
```


### 1. Designate ACS variables to be pulled

* The first step in pulling a list of ACS variables is to desigate the list of variables to be pulled. There is a built in list of ACS variables (ACSColumns.csv) that is downloaded with the GecoderACMT.R code. The interpolation strategy (to sum or not) for that built in list of ACS variables is written out in the GeocoderACMT.R code as well. We can modify this list of variables and the interpolation lists for the variables that we are interested in. Make sure the list of Inspace ACS variables to pull is uploaded into your workspace/Inspace folder, then proceed as follows.

* Create a list of the variables of interest. *

```{r set acs vars}
acs_columns_inspace<-read.csv('Inspace/ACSColumns_inspace.csv') %>% dplyr::select(-X)
acs_columns_inspace$acs_variable_name_to_interpolate_by_sum_boolean_mapping<-acs_columns_inspace$interpolation

#overwrite current ACSColumns.csv file
write.csv(acs_columns_inspace, 'ACMT/ACSColumns.csv')

```

* If you are using a year prior to 2015, check with the InSpace team so that they can verify that those same variable columns codes apply to the same variables for that year. 

### 2. Designate settings (radius, year, count names, proportion names) for ACS variable pull

  Next, we will create count and proportion names from the list of ACS variables that we just created, and designate those named count and proportion variables for our 'names_of_variables_to_get' vector. We also need to designate the codes_of_acs_variables_to_get vector from our ACS variable dataframe and interpolation strategy. In addition, we can desigate the radius (or radii) and year for pulling variables. 

```{r designate names and settings}
##create 'count' versions of each variable name and 'proportion' versions for each #ACS variable where applicable
acs_count_names<-paste(acs_columns_inspace$var_name, "count", sep="_")
if (length(acs_columns_inspace$var_name[acs_columns_inspace$universe_col != ""]) == 0) {   # prevent having something that is exactly "_proportion"
  acs_proportion_names <- character(0)
} else {
  acs_proportion_names <- paste(acs_columns_inspace$var_name[acs_columns_inspace$universe_col !=''], "proportion", sep="_")   # only non-universal variables have proportions
}

#Designate interpolation strategy for variables
acs_variable_name_to_interpolate_by_sum_boolean_mapping<-acs_columns_inspace$interpolation
names(acs_variable_name_to_interpolate_by_sum_boolean_mapping)<-acs_columns_inspace$acs_col

#Set the list of variable codes, the list of variable names, the radius, and the year for the data you want pulled
codes_of_acs_variables_to_get<-acs_columns_inspace$acs_col
names_of_variables_to_get<-c(acs_count_names, acs_proportion_names)
radius_vector <- c(500, 1000, 5000)#set the radius for the area of interest
years <- c(2017, 2019) #set the year(s) for the data of interest

```


### 3. Create additional columns in the dataset for each variable to be pulled

In order to directly add the measures to your dataset, we first need to add columns for each variable and rows for years/radii
  

```{r process geocoded data}
dataset_geocode<-read.csv('Inspace/dataset_geocoded.csv') %>% dplyr::select(id, lat, long)

#dataset_geocode<-dataset_geocode%>%
#  rename(lat=latitude)%>%
#  select(id, lat, long)

head(dataset_geocode)
```

Your imported dataset should be formatted as follows, with the following column names: 

| id  |  lat      | long        | 
|-----|-----------|-------------|
| 01  | 47.568922 | -122.306422 |
| 02  | 47.632264 | -122.314978 |
| 03  | 47.634820 | -122.292769 |
| ..  | ......... | ........... |

To this dataset, we will add columsn for each variable to be pulled, and rows for each year/radii: 

```{r process dataframe}
dataset_acs<-add_rows_columns(dataset=dataset_geocode, variable_list=names_of_variables_to_get, radius_vector=radius_vector, years=years)
head(dataset_acs)
```

### 4. Run the ACMT Loop to pull ACS data and interpolate each measures for each lat/long

Finally, we will run this loop to pull the ACS Measures for each lat/long and each year/radii in our dataset. This loop is set up so it can be interrupted, and restart from this step. If the loop is interrupted and you close R, be sure to save your dataset_acs into your Inspace folder before shutting down. 

If returning to this data pull after shutting down R, you will need to reimport your dataset_acs, and rerun the code in the introduction through step 2 (NOT STEP 3). 

```{r acmt loop, warnings='hide', results='hide'}

## to reimport your dataset_acs after interruping the process (don't forget to run code in the introduction through step 2)
dataset_acs<-read.csv('Inspace/dataset_acs.csv')%>%mutate(X=NULL)

#run loop to pull variables
for(year in 1:length(years)){
  year<-years[year]
  print(year)
for(radius in 1:length(radius_vector)){
  #radius<-1 #for testing
radius<-radius_vector[radius]
#dataset_radius<-dataset_acs
print(radius)

for(id in 1:nrow(dataset_acs)) {
   tryCatch({if(!is.na(dataset_acs[,1][id])) next #skip the row if the data is already there
  print(id) #print the number to keep track of progress
  latitude<-dataset_acs[dataset_acs$radius==radius & dataset_acs$year==year,]$lat[id] #set lat
  longitude<-dataset_acs[dataset_acs$radius==radius & dataset_acs$year==year,]$long[id] #set long
  
  environmental_measures<-get_acmt_standard_array(long=longitude, lat=latitude, radius_meters = radius, year=year, codes_of_acs_variables_to_get = codes_of_acs_variables_to_get, set_var_list=TRUE) #pull measures for given lat & long
 
#      for(name_of_variable in names_of_variables_to_get){ #for each measures, get the value and put it into the column of the same name
#     value_of_variable <- environmental_measures[environmental_measures$names == name_of_variable,]$values  
#     dataset_radius[[name_of_variable]][id]<-value_of_variable
#  }

 for (name_of_variable in names_of_variables_to_get) {
        dataset_acs[dataset_acs$radius==radius & dataset_acs$year==year,][[name_of_variable]][id]<- environmental_measures[environmental_measures$names == name_of_variable, ]$values  
 }
  },error=function(e){cat("ERROR :", conditionMessage(e), "\n")}) #this will print any error messages
}
}
}  

#### Note: this loop is set up to fill in your dataset as it runs, so if the loop is interrupted, you can re-run this chunk only, and it will skip over data that is has already pulled. ####

getwd()

write.csv(dataset_acs, 'Inspace/dataset_acs.csv')

```


### 5. Next, the county GEOID for each lat/long location

```{r pull county GEOID}
#download cbsa shapefile
#cbsa<-core_based_statistical_areas(year=2019)

#lat-long shapefile with 
latlong_sf<-st_as_sf(dataset_acs%>%dplyr::select(id, lat, long), coords=c('long', 'lat'), crs=4269)
counties_sf<-st_transform(counties, crs=4269)
#cbsa_sf<-st_as_sf(cbsa, crs=4269)
#intersected_msa<-st_intersection(latlong_sf, cbsa_sf)
intersected_county<-st_intersection(latlong_sf, counties_sf)

#msa_dataset <- merge(dataset_acs %>%dplyr::select(id), intersected_msa, by='id', all=TRUE)%>% dplyr::select(id, #msa_geoid=GEOID)
county_dataset<-merge(dataset_acs%>%dplyr::select(id), intersected_county, by='id', all=TRUE)%>%dplyr::select(id, county_geoid=GEOID, -geometry)

#dataset_acs_full<-merge(dataset_acs_full, msa_dataset, by='id', all.x=TRUE)
dataset_acs<-merge(dataset_acs, county_dataset, by='id', all.x=TRUE)

```

### 6. Finally export your dataset

```{r export}
dataset_acs<-dataset_acs %>% dplyr::select(-lat, -long)
write.csv(dataset_acs, 'Inspace/dataset_acs.csv')

```

