---
title: "Gentrification Data Pull"
author: "Amy Youngbloom & Najma Abdi"
date: "10/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
```

## Introduction

This measure of gentrification comes from the Urban Health Collaborative (UHC) at Drexel University and conceptualize gentrification as a neighborhood change process that occurs over a period of time. The UHC measure identifies census tracts that have gentrified using a two-step process: 1) Identify census tracts that are eligible to gentrify at a specific baseline 2) Of those that were eligible, identify the tracts that show evidence of moderate or intense gentrification between baseline and a follow-up year. This results in four categories of gentrification: ineligible to gentrify (9), eligible to gentrify but did not gentrify (0), evidence of gentrification (1), intense gentrification (2). 

## 1. Set up ACMT & Import dataset

First, we need to load the ACMT packages by calling the following function
```{r}
setwd('~/workspace')
source("setup-acmt.R")
```

Next, import your geocoded dataset

```{r}
setwd("/home/rstudio/workspace/Inspace")
dataset_geocoded<-read.csv('dataset_geocoded.csv')
```

## 2. Pull tracts for each participant

```{r pull tracts data}
#designate state
state<-'WA'

#download tract shapefile
tracts.2010<-tracts(year=2010, state=state)

#create lat-long shapefile, msa shapefile, and state shapefiles with the same projection
latlong_sf<-st_as_sf(dataset_geocoded, coords=c('long', 'lat'), crs=4269)
tract_2010_sf<-st_as_sf(tracts.2010, crs=4269)

#find intersections of lat/long with tracts
intersected_tracts<-st_intersection(latlong_sf, tract_2010_sf)%>%dplyr::select(id, GEOID10)%>%st_drop_geometry()
head(intersected_tracts)

```

### 3. Download Gentrification data

```{r import price parity data}

#set URL to download from github
genturl<-'https://raw.githubusercontent.com/aybloom/inspace/main/docs/Inspace/raw_gentrification.csv'

download.file(url = genturl, destfile = "~/workspace/external_data/downloaded_gentrification.csv")


```

### 4. Process Gentrification data

```{r process data}
raw_gent_data<-read.csv('~/workspace/external_data/downloaded_gentrification.csv')

processed_gentrification<- raw_gent_data %>%
  rename(GEOID10=geoid10)%>%
  mutate(GEOID10=ifelse(GEOID10<10000000000, as.character(paste0('0', as.character(GEOID10), "")), as.character(GEOID10))) #convert to GEOID to character for joining data, need to add an extra 0 in front for some values

head(processed_gentrification)


```

### 5. Pull gentrification data for participants


```{r pull gent data}

dataset_gentrification<-merge(intersected_tracts, processed_gentrification, by='GEOID10', all.x=TRUE)
head(dataset_gentrification)

```



### 4. Export your dataset

```{r export data}

write.csv(dataset_gentrification, '~/workspace/Inspace/dataset_gentrification.csv')

```
