---
title: "Checking addresses with geocoder & mapping tool"
author: "Amy Youngbloom"
date: "10/27/2022"
output: html_document
---

```{r setup, include=FALSE}
#knit this file with the working directory of "workspace/" instead of "workspace/examples"
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, eval=TRUE)
knitr::opts_knit$set(root.dir = '/home/rstudio/workspace/')
```

```{r load packages & the ACMT}
library(tidycensus)
library(tidyverse)
library(dplyr)
library(janitor)
library(reshape2)
#install.packages('ggmap')
library(ggmap)
source("GeocoderACMT.R")
```
If the above code results in an error, you may need to remove the '#' before install.packages('ggmap'). 

## InSpace – Geocoding Protocol

	Ensuring the most precise geocoding of participant addresses will enable you to pull the most accurate measures of the environmental context. While the ACMT has a built-in geocoder, we are also relying on our partners to verify that the geocodes produced by the ACMT accurately represent the location of the participant. 
	
	The process of verifying participant locations is an iterative process of mapping geocodes, updating addresses, re-mapping, and noting where location are inaccurate, even after address updates. This process utilizes the ACMT map checking and geocoder functions to produce maps, geocodes, and ratings of the geocodes. 

## 1. Import address data & combine address fields if necessary

* You will first need to upload your datatset into your Inspace folder. 
* Next, you will update the code below with the name of your dataset


```{r import dataset}
dataset<-read.csv('Inspace/dataset_example.csv') %>% dplyr::select(-X) ##update with the name of your dataset
colnames(dataset) 
```

Look at the column names in your dataset. You should have an id field titled 'id'. 

* If your address field is already in one column, you can skip the chunk of code below. 
* If your address fields are in separate columns, (i.e., street, city, state, zip), you can use the code below to combine your addresses into one field.

```{r combine address fields}
dataset<-dataset %>%
  #rename(id=ID, address=oldAddressName) %>% # use this line to update your ID column name
  mutate(address=paste(street, city, state, zip, sep=', ')) %>% ##Ensure your dataset has these column names
  dplyr::select(id, address)

head(dataset)
```

## 2. Write mapping functions

We can get a map showing the results of the geocode function given an address. In the title of the map, the address and the rate are listed. The point with coordinates equal to the calculated longtitude and latitude is plotted. This is a good way to verify the accuracy of geocoding. First, we create the functions:

```{r check geocode function}
check_geocode <- function(lat, long, address, rate, id, z=16, side_len=0.007){
  bbox <- c(left = long - side_len, bottom = lat - side_len, right = long + side_len, top = lat + side_len)
  map <- ggmap(get_stamenmap(bbox, zoom = z)) +
    geom_point(x = c(long), y = c(lat), size = 3) +
    ggtitle(paste(paste('ID = ', id, sep=''), address, paste("Rate =", rate, sep = ""), sep='\n'))
  return(map)
}
check_geocode_for_address <- function(address, id){
  location <- geocode(address = address)
  map <- check_geocode(lat = location$latitude, long = location$longitude, address = address, rate = location$rating, id=id)
  return(map)
}


```

## 3. Run Mapping function loop to map all addresses

Next, we will geocode and print maps for all individuals in order to check the geocoded locations.

This loop will create a pdf file with a map printed for each individual in your dataset as well as a csv file with ratings, address, and a notes columns. You'll find these documents in your Inspace folder. Export them and use them to review the maps. 


```{r geocode and get ratings}
#dataset<-read.csv('Inspace/dataset_address_check.csv')%>%dplyr::select(id, address)

## geocode addresses
#create lat and long columns
dataset_geocode<-dataset %>%
  mutate(lat=NA, 
         long=NA, 
         rating=NA)
#Geocoding loop
for (i in 1:nrow(dataset_geocode)) {
  if(!is.na(dataset_geocode$lat[i])) next #skip already geocoded
  if(is.na(dataset_geocode$address[i])) next #skip NA address values
  print(i) #print the number to track progress
address<-dataset_geocode$address
  lat_long<-tryCatch({geocode(address[i])}, 
                     error=function(x){
                       lat_long<-list(latitude=NA, longitude=NA, rating=NA)
                       })
  dataset_geocode$lat[i]<-lat_long$latitude # add latitude to dataset_geocode
  dataset_geocode$long[i] <-lat_long$longitude # add longitude to dataset_geocode
  dataset_geocode$rating[i]<-lat_long$rating # add rating to the dataset_geocode
}
dataset_geocode<-dataset_geocode %>% arrange(-rating)
write.csv(dataset_geocode %>% dplyr::select(id, address, rating)%>%mutate(map_check_notes=""), 'Inspace/dataset_address_check.csv') ##writes a csv file with address and ratings and a column for update notes;

ratings.table<-dataset_geocode %>%
  mutate(rating_group=ifelse(rating==0, '0', ifelse(rating>0 & rating<10, '0 to 9', ifelse(rating>9 & rating < 21, '10-20', ifelse(rating>20, '20 or higher', NA)))))%>%group_by(rating_group)%>%summarise(n=n())%>%mutate(percent=round(n/sum(n), 2)) %>%
  dplyr::select(rating_group, n, percent)

ratings.table
```

## take a screenshot of the rating table above and save. 

Next, we will print a pdf file with the initial maps for all of your geocoded addresses. 

```{r print maps, eval=FALSE}
pdf('Inspace/geocode_checker_maps.pdf', height=10, width=10)
for(i in 1:nrow(dataset_geocode)){
  print(i) #print the number to track progress
  #set values for function
  lat<-dataset_geocode$lat[i]
  long<-dataset_geocode$long[i]
  address<-dataset_geocode$address[i]
  rating<-dataset_geocode$rating[i]
  id<-dataset_geocode$id[i]
  
map<-tryCatch({check_geocode(lat=lat, long=long, address=address, rate=rating, id=id)}, error=function(x){map<-NA})

 print(map)
}

```

## 4. Check the mapped geocodes

  * Open up the pdf of maps file that should now be in your Inspace folder (you may need to reload your Inspace folder), called, 'geocode_checker_maps'. 
  * Export the csv file (called dataset_address_check) and open it in excel. You will be working off of this file check geolocations and make notes about addresses that have high ratings, but look accurate.
  
### For each address with rating >0, 

For each address with a geocoding rating >0, and whose map is either blank, or appears inaccurate, take the following steps. After each step, you can use the code below, updating the i <- ## with the id number you are updating, and the update.address<- '' with the updated address text. 

  1. Update the address to write out any 'Ave' to 'Avenue', 'S' to 'South', etc. 
  2. Check the zip code for the given city and/or cross street. 
  3. Increase/Decrease the level of detail by changing the 'z' value (higher numbers here give more detailed maps, but may not be available for more rural areas - you can lower the number to 14 or 15 if the map is blank)
  4. Increase/Decrease the level of zoom by changing the 'side_len' value (higher numbers give a larger map, but may take longer to load (I recommend not going higher than 0.05))
  5. If you are unsure if the location is correct or not, and the rating is >0, 
  
    * update the code below wth the id number of the individuals whose address you are looking to update, and paste the entire new address in the code below to update: 
  
```{r update address, fig.height=10, fig.width=10}

## enter the id number for the participant whose address you are updating here: 
i<- 5 #
#i<- 30
## enter the updated address text for the individual here  
update.address<- '3013 S Mount Baker Blvd., Seattle, WA, 98144' 

#update the address & reprint the map
dataset_geocode$address[dataset_geocode$id==i]<-update.address
dataset_geocode$address[dataset_geocode$id==i]

## get updated geocodes
address<-dataset_geocode$address[dataset_geocode$id==i]
lat_long<-tryCatch({geocode(address)}, 
                     error=function(x){
                       lat_long<-list(latitude=NA, longitude=NA, rating=NA)
                       })
  dataset_geocode$lat[dataset_geocode$id==i]<-lat_long$latitude # add latitude to dataset_geocode
  dataset_geocode$long[dataset_geocode$id==i] <-lat_long$longitude # add longitude to dataset_geocode
  dataset_geocode$rating[dataset_geocode$id==i]<-lat_long$rating # add rating to the dataset_geocode


##reprint the mapped address
check_geocode(lat=dataset_geocode$lat[dataset_geocode$id==i], long=dataset_geocode$long[dataset_geocode$id==i], address=dataset_geocode$address[dataset_geocode$id==i], rate=dataset_geocode$rating[dataset_geocode$id==i], id=i, 
              z=17, ##higher numbers here give more detailed maps, but may not be available for more rural areas - you can lower the number to 14 or 15 if the map is blank. 
              side_len=0.005 # higher numbers give a larger map, but may take longer to load (I recommend not going higher than 0.05)
              )

## if the reprinted map still looks incorrect, make additional edits to the address and rerun the code. 

```

 
*NOTE: Do not put any participant addresses into google maps. If you need to check the address, you can put the street name into google maps, but do not include the house number*

##export dataset to Inspace folder

```{r export dataset}

write.csv(dataset_geocode %>% dplyr::select(id, lat, long), 'Inspace/dataset_geocoded.csv')

```
  
  