---
title: "mRFEI for Inspace"
author: "Amy Youngbloom"
date: "4/15/2022"
output: html_document
---

```{r setup, include=FALSE}
# knit this file with the working directory of "workspace/" instead of "workspace/examples"
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, eval=FALSE)
knitr::opts_knit$set(root.dir = '~/workspace')
```

## The Modified Food Retail Environment Index

  - The CDC’s modified retail food environment index provides census tract level estimates the ratio of healthy food outlets to total food outlets.  We will use area-weighted interpolation of tract-level estimates to construct buffer-specific estimates of the following measures. 
  
  - The mRFEI settings are built into the ACMT external data file_loader, external_data_presents, and external_data_file_downloader R code files. 
  
```{r}
##Ensure workspace is set to 'workspace' folder, and run the following R files: 
setwd('~/workspace')
source('external_data-file_downloader_and_processor.R')
source('external_data-file_loader.R')
source('external_data-presets.R')
source('setup-acmt.R')

```

The mRFEI dataset is built into the ACMT, making accessing this data easy. This guide provides steps on pulling mRFEI data for your geocoded dataset

## 1. First we will import your geocoded dataset and add columns for each measure to be pulled

```{r import geocoded data}
setwd("/home/rstudio/workspace/Inspace")
dataset_geocode<-read.csv('dataset_geocoded.csv')

mrfei_vars<-c('total_pop_count', # total population measure from acs
'mRFEI')

#create dataset with a column for each variable at each radius level
var.cols<-data.frame(matrix(nrow=nrow(dataset_geocode), ncol=(length(mrfei_vars)))) #create dataset of columns
colnames(var.cols)<-mrfei_vars #name the columns

dataset_mrfei<-cbind(var.cols, dataset_geocode)

```
 
## 2. Designate settings

Next we set the year and radius for the ACMT. We need to include at least 1 ACS variable, otherwise the function will try to pull all of the ACS variables in the ACS Columns file (which will take a while)
  
```{r designate acmt settings, results='hide'}

#Set the list of variable codes, the list of variable names, the radius, and the year for the data you want pulled
codes_of_acs_variables_to_get<-'B01001_001'
names_of_variables_to_get<-mrfei_vars
radius_vector <- c(500, 1000, 5000)#set the radius for the area of interest
year <- c(2019) #set the year for the data of interest

# set up mrfei settings:
external_data_name_to_info_list <- list(
  mrfei=external_data_presets_mrfei
)

```

## 3. Run loop to pull mRFEI data

```{r data pull loop, results='hide'}

## Create loop for pulling mRFEI values for each address in the dataset
setwd('/home/rstudio/workspace/')

#run loop to pull variables
for(radius in 1:length(radius_vector)){
  #radius<-1 #for testing
radius<-radius_vector[radius]
print(radius)
dataset_radius<-dataset_mrfei

for(address in 1:nrow(dataset_radius)) {
   tryCatch({if(!is.na(dataset_radius[,1][address])) next #skip the row if the data is already there
  if(!is.na(dataset_radius[,1][address])) next #skip the row if the data is already there
  print(address) #print the number to keep track of progress
  latitude<-dataset_radius$lat[address] #set lat
  longitude<-dataset_radius$long[address] #set long
  
  environmental_measures<-get_acmt_standard_array(long=longitude, lat=latitude, radius_meters = radius, year=year, codes_of_acs_variables_to_get = codes_of_acs_variables_to_get, external_data_name_to_info_list=external_data_name_to_info_list, fill_missing_GEOID_with_zero = TRUE) #pull measures for given lat & long
 
 
for (name_of_variable in names_of_variables_to_get) {
        dataset_radius[[name_of_variable]][address] <- environmental_measures[environmental_measures$names == name_of_variable, ]$values  
 }},error=function(e){cat("ERROR :", conditionMessage(e), "\n")}) #this will print any error messages
}

if(radius==500) {
dataset_mrfei2<-dataset_radius
dataset_mrfei2$radius<-500}

if(radius>500){
dataset_radius$radius<-radius
dataset_mrfei2<-rbind(dataset_mrfei2, dataset_radius)
}

}

dataset_mrfei2<-dataset_mrfei2%>%
  mutate(year=year)%>% # set year of data pull
  dplyr::select(-lat, -long, id, radius, year, everything()) # reorder variables

```

## 4. Export dataset

```{r export data}
setwd("/home/rstudio/workspace/Inspace")
write.csv(dataset_mrfei2, 'dataset_mrfei.csv')


```

