---
title: "National Land Cover Database"
author: "Amy Youngbloom"
date: "4/13/2022"
output: html_document
---

```{r setup, include=FALSE}

# knit this file with the working directory of "workspace/" instead of "workspace/examples"
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, eval=FALSE)
knitr::opts_knit$set(root.dir = '~/workspace')
library(tidyverse)
```

```{r load acmt}
setwd('~/workspace')
source('setup-acmt.R')
```

### Introduction

  The National Land Cover Database uses satelitte imagery to create pixel-by-pixes maps of land cover features. Areas of 30-meters are categorized into land cover classes including forest, low-, medium-, and high-intensity developed, and more. For more information about the NLCD, visit this link: https://www.usgs.gov/news/technical-announcement/new-land-cover-maps-capture-nearly-two-decades-change-across-us
  

## 1 Download NLCD data

  - To download NLCD data, navigate in another window to the MRLC.gov NLCD viewer site: https://www.mrlc.gov/viewer/
  - In the viewer, zoom into the state of interest. 
  - Next click the 'download' button in the top left of the map, which will allow you to select a box on the map that includes your area of interest
  - Click on the map where the top left corner of your box will start, then bring your curser to where the bottom right of your box will be and click again. A partially transparent box will cover your area. 
  - A Data Download section will open on the right side of the screen where you can enter your email, select 'Land Cover', then click download.
  - You should receive an email shortly with the data download.

```{r github, echo=FALSE, fig.cap="Download ACMT source code from github", out.width = '50%'}
knitr::include_graphics(paste0(getwd(),"/Inspace/nlcd-image.png"))
```

  - Once data is downloaded, extract the data and upload into workspace folder. You should see a landcover tiff file for: 2001, 2004, 2006, 2008, 2011, 2013, 2016, and 2019. 
  
  - Depending on the year you are looking for, identify the correct Tiff file to import into R. In this example, we will use the 2019 data. Rename your file to match the file name in the raster() function below (you may want to update the year in code below if you are using a different year)

# 2 Import NLCD data into R

```{r import raster data & geocoded dataset, results='hide'}
##update the data file name to match the file you've imported into workspace:
setwd('~/workspace/Inspace')
##set year(s) of interest
year<-2016

##pull data for given year
if(year>2017){
nlcd_data<- raster('NLCD_2019_Land_Cover.tiff')
}
if(year<2018 & year>2014){
nlcd_data<-raster('NLCD_2016_Land_Cover.tiff')
}
if(year<2015){
nlcd_data<-raster('NLCD_2013_Land_Cover.tiff')  
}

##Import geocoded dataset
dataset_geocode<-read.csv('dataset_geocoded.csv')
dataset_nlcd<-dataset_geocode

```

## 3 Create buffer areas around each data point:

- First, we'll re-set these two functions (they are also written into the ACMTGeocoder)

```{r set up functions, results='hide'}

get_projection_for_lat_long <- function(long, lat) {
  point <- st_sfc(st_point(c(long, lat)), crs=4326)
  state_plane_zones %>%
    filter(st_contains(state_plane_zones, point, sparse = F) == 1) %>%
    left_join(state_proj, by="ZONE") %>%
  {.} -> selected_zone
  if (nrow(selected_zone) == 0) {
    search_factor <- 1
    while (nrow(selected_zone) == 0) {
      point <- st_sfc(st_point(c(long+runif(1, -0.1*search_factor, 0.1*search_factor),
                                 lat+runif(1, -0.1*search_factor, 0.1*search_factor))), crs=4326)
      state_plane_zones %>%
        filter(st_contains(state_plane_zones, point, sparse = F) == 1) %>%
        left_join(state_proj, by="ZONE") %>%
      {.} -> selected_zone
      search_factor <- search_factor + 1
    }
  }
  return(selected_zone$proj4_string)
}

get_point_buffer_for_lat_long <- function(long, lat, radius_meters) {
  proj4_string <- get_projection_for_lat_long(long, lat)
  point <- st_sfc(st_point(c(long, lat)), crs=4326)
  point_projected <- st_transform(point, proj4_string)
  radius <- set_units(radius_meters, "meters")
  point_buffer <- st_buffer(point_projected, dist=radius)
  point_buffer <- st_transform(point_buffer, crs=4326)
  point_buffer<-st_sf(point_buffer)
}

```

```{r loop to get buffers, results='hide'}

#set radius: 
radius_vector<-c(500, 1000, 5000)

for(radius in radius_vector){
  radius=radius
  print(radius)
  dataset_radius<-dataset_nlcd
for (id in 1:nrow(dataset_radius)){
print(id)
    tryCatch({
#set long and lat
long<-dataset_radius$long[id]
lat<-dataset_radius$lat[id]
#create buffer shapefile
buffer1<-get_point_buffer_for_lat_long(long, lat, radius_meters=radius)

##extract NLCD data for buffer
nlcd_extract<-raster::extract(nlcd_data, buffer1)
#calculate proportion of each type of land cover, convert prop.table to dataframe
prop.nlcd<-lapply(nlcd_extract, FUN=function(x) {prop.table(table(x))})
prop.tb<-data.frame(prop.nlcd)
#tryCatch({
nlcd.table<-spread(prop.tb, x, Freq)
nlcd.table<-nlcd.table %>%
  mutate(id = dataset_radius$id[id], 
         radius=radius)

if (id == 1){
  nlcd.data.full<-nlcd.table
  }
if(id>1){
  nlcd.data.full<-plyr::rbind.fill(nlcd.table, nlcd.data.full)
}
  },error=function(e){cat("ERROR :", conditionMessage(e), "\n")}) #this will print any error messages})

}
if(radius==500){
  dataset_nlcd2<-nlcd.data.full
}
  if(radius>500){
    dataset_nlcd2<-plyr::rbind.fill(dataset_nlcd2, nlcd.data.full)
  }
  }

dataset_nlcd2<-dataset_nlcd2%>%
  mutate_all(~replace(., is.na(.), 0))%>%   #replace NA values with 0
  dplyr::select(id, radius, everything())

head(dataset_nlcd2)
```

Proportion of buffer that is water
  - Proportion of buffer that is developed open space
  - Proportion of buffer that is developed low intensity
  - Proportion of buffer that is developed medium intensity
  - Proportion of buffer that is developed high intensity
  - Proportion of buffer that is any undeveloped

```{r update nlcd columns}

##udpate colnames -- you may need to # out columns that don't apply to your dataset
dataset_nlcd2 <- dataset_nlcd2 %>%
  rename('open_water'='11', 
  'develop_openspace'='21', 
  'develop_lowintense'='22', 
  'develop_medintense'='23', 
  'develop_highintense'='24', 
  'barren_land'='31', 
  'decid_forest'='41', 
  'evergreen_forest'='42', 
  'mixed_forest'='43', 
  #'dwarf_scrub'='51', 
  'shrub_scrub'='52', 
  'grassland'='71', 
  #'sedge_herb'='72', 
  #'lichen'='73', 
  #'moss'='74', 
  'pasture'='81',
  'cultivate_crops'='82', 
  'wood_wetlands'='90', 
  'herb_wetlands'='95') %>%
  rowwise()%>%
  mutate(undeveloped = sum(c(barren_land, decid_forest, evergreen_forest, mixed_forest, 
                           #dwarf_scrub, 
                           shrub_scrub, grassland, 
                           #sedge_herb, lichen, moss, 
                           pasture, 
                           cultivate_crops, 
                           wood_wetlands, herb_wetlands), na.rm=TRUE))%>% mutate(year=year)
 
head(dataset_nlcd2) 
```

## 4. Exporting datasets

### for multiple years

If you are pulling data for multiple years, save the dataset, then return to step 2, change the year, and re-pull the data for the additional year.

```{r set year}
assign(paste('dataset_nlcd2', year, sep='_'), dataset_nlcd2)

```

### export dataset - for 2 years: 

If you've pulled data for 2 years, run this step to combine datasets (you may need to update the dataset names if your years differ)

```{r combine datasets}

dataset_nlcd_full<-rbind(dataset_nlcd2_2015, dataset_nlcd2_2019)

write.csv(dataset_nlcd_full, '~/workspace/Inspace/dataset_nlcd.csv')
```


### export dataset -- for only 1 year:

```{r export datset}
setwd('~/workspace/Inspace')
write.csv(dataset_nlcd2, 'dataset_nlcd.csv')

```
