---
title: "External Data - PLACES"
author: "Amy Youngbloom"
date: "4/8/2022"
output: html_document
---

```{r setup, include=FALSE}
#knit this file with the working directory of "workspace/" instead of "workspace/examples"
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, eval=FALSE)
knitr::opts_knit$set(root.dir = '/home/rstudio/workspace/')
```
## Introduction: CDC PLACES Data

https://www.cdc.gov/places/about/index.html

The Places dataset provides census tract level estimates of population behaviors taken from BRFSS responses.  As with the ACS data, we will use area-weighted interpolation of tract-level estimates to construct buffer-specific estimates of several measures. PLACES was an expasion of the 500 Cities project (https://www.cdc.gov/places/about/500-cities-2016-2019/index.html) and began in 2020, utilizing BRFSS data from 2017 and 2018. Since then the 2021 BRFSS has also been released and utilized 2019 BRFSS data. Below are instructions for pulling the 2020 or the 2021 PLACES data or the 2016 - 2019 500 Cities datasets. 

For this dataset, you will pull the dataset using 2015 BRFSS data (cities2017) as well as the dataset that includes the year you began enrollment.  Because the data estimates across 2 years, choose the dataset that makes the most sense with the timing of your enrollment. For example, if enrollment began in late 2017 and continued into 2018, you would use the places 2020 dataset. If enrollment took place in the first part of 2017, you would use the cities 2019 dataset. 

### 1. Setup ACMT

```{r setup acmt}
source('setup-acmt.R')
source('Inspace/inspace_external_data_functions.R')
```

### 2. Designate settings for data pull

```{r designate settings}
years<-c(2018) #set year of data pulls: only 2017 and 2018 are available for all census tracts -- pull the year closest to enrollment
radius_vector<-c(500,1000, 5000) #set radii
variable_list<-names(places_variable_name_to_interpolate_by_sum_boolean_mapping) ## set variables -- should be designated in the external_data_presets
names_of_variables_to_get<-variable_list
external_data_name_to_info_list <- list(places=external_data_presets_places) ## update with presets for external dataset (be sure presets are designated in the external_data-presets.R)
```

### 3. Import dataset and process

```{r}
dataset_geocoded<-read.csv('~/workspace/Inspace/dataset_geocoded.csv')%>%dplyr::select(id, lat, long)
dataset_places<-add_rows_columns(dataset_geocoded, variable_list=variable_list, years=years, radius_vector=radius_vector)

### get states of interest from datatset
states_sf <- st_transform( us_states( map_date = NULL, resolution = c("low", "high"), states = NULL), 4326)
points_sf = st_as_sf(dataset_geocoded, coords = c("long", "lat"), crs = 4326, agr = "constant")
states <- as.data.frame( st_join(points_sf, states_sf, join = st_intersects) ) %>% dplyr::select(state_abbr, -geometry)%>%unique()%>% as.list()


```

### 4. Run loop to pull measures for each lat/long

```{r acmt loop, results='hide'}

#run loop to pull variables
for(i in 1:length(years)){
##remove downloaded data befor each year so it will be replaced with appropriate dataset -- if the files don't exist, will print and error and continue; 
file.remove('~/workspace/external_data/processed_places.csv')
#file.remove('home/rstudio/workspace/external_data/downloaded_places.csv')
year<-years[i]
print(year)
for(radius in 1:length(radius_vector)){
radius<-radius_vector[radius]
print(radius)

for(address in 1:nrow(dataset_places[dataset_places$radius==radius & dataset_places$year==year,])) {
   tryCatch({if(!is.na(dataset_places[dataset_places$radius==radius & dataset_places$year==year,][,1][address])) next #skip the row if the data is already there
print(address) #print the number to keep track of progress
  latitude<-dataset_places[dataset_places$radius==radius & dataset_places$year==year,]$lat[address] #set lat
  longitude<-dataset_places[dataset_places$radius==radius & dataset_places$year==year,]$long[address] #set long
  
##pull measures for each lat/long 
  environmental_measures<-get_acmt_standard_array(long=longitude, lat=latitude, radius_meters = radius, year=year, codes_of_acs_variables_to_get = NULL, external_data_name_to_info_list=external_data_name_to_info_list, fill_missing_GEOID_with_zero = FALSE) #pull measures for given lat & long

   for (name_of_variable in names_of_variables_to_get) {
        dataset_places[dataset_places$radius==radius & dataset_places$year==year,][[name_of_variable]][address] <- environmental_measures[environmental_measures$names == name_of_variable, ]$values 
        
 }
  },error=function(e){cat("ERROR :", conditionMessage(e), "\n")}) #this will print any error messages
}
}
}

#### Notes: this loop is set up to fill in your dataset as it runs. If it is interrupted, no problem. You can re-run this chunk only, and it will skip over data that is has already pulled. ####
```
  
### 6 Export your data 

  Remove lat/long values and export your data

```{r export data}

dataset_places<-dataset_places %>% dplyr::select(everything(), -lat, -long)
write.csv(datset_places, 'Inspace/dataset_places.csv')

```


