---
title: "Sidewalk View - Data Pull"
author: "Amy Youngbloom"
date: "5/17/2022"
output: html_document
---

```{r setup, include=FALSE}
# knit this file with the working directory of "workspace/" instead of "workspace/examples"
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, eval=FALSE)
knitr::opts_knit$set(root.dir = '~/workspace')
```


## Sidewalk View 

     Quynh Nguyen at University of Maryland has developed a national sidewalk presence dataset by applying machine learning to Google Street View images to produce tract-level estimates of sidewalk presence/panoramic image by tract. Measures included percent of images for a given census tract that has crosswalks, the percent of images for a given census tract with a sidewalk We will use area-weighted interpolation to construct buffer specific measures of sidewalk presence. 
     
     
## Import dataset

  The first step is to download the sidwalk data and save it in the external_data folder as 'downloaded_sidewalk.csv'
  
### 1. Process sidewalk view data

Next we will write a function to process the sidewalk view dataset into a 'ACMT-friendly' (long) format. 

```{r process data}
process_sidewalk<-function() {
  raw_sidewalk<-read.csv('external_data/downloaded_sidewalk.csv')

processed_dataframe<-raw_sidewalk %>%
  dplyr::select(censustract, total_num, total_crosswalk, total_sidewalk) %>%
    melt(id='censustract')%>%
    rename(GEOID=censustract, estimate=value) %>%
    mutate(GEOID=as.character(GEOID)) %>%
    mutate(GEOID=ifelse(nchar(GEOID)<11, paste0('0', GEOID), GEOID)) #convert to GEOID to character for joining data, need to add an extra 0 in front for some values

processed_dataframe$estimate[is.na(processed_dataframe$estimate)]<-0 #impute NA with 0 values
  
write_csv(processed_dataframe, 'external_data/processed_sidewalk.csv')
}

```

### 3. Designate settings for ACMT

Next we need to designate settings for the ACMT, including the vector_of_expected_dowloaded_file_name, the process_file, the geoid_type, and the variable_name_to_interpolate_by_sum_boolean_mapping

```{r, results='hide'}
setwd('~/workspace')
source("setup-acmt.R")

#set boolean mapping for interpolation
variable_name_to_interpolate_by_sum_boolean_mapping = c(TRUE, TRUE, TRUE, FALSE, FALSE)

names(variable_name_to_interpolate_by_sum_boolean_mapping) = c('total_num', 'total_crosswalk', 'total_sidewalk', 'prop_crosswalk', 'prop_sidewalk')

#designate settings for external data pull
external_data_name_to_info_list <- list(
        sidewalk=list(vector_of_expected_downloaded_file_name=c("downloaded_sidewalk.csv"),
                   download_file=NULL,
                   process_file=process_sidewalk,
                   geoid_type="Census Tract",
                   variable_name_to_interpolate_by_sum_boolean_mapping=variable_name_to_interpolate_by_sum_boolean_mapping 
        )
)

```


## 4. Import geocoded dataset & add column names for measures to pull
- Next, we create empty dataset columns for each variable to be pulled. 

```{r set variable columns, results='hide'}
setwd('~/workspace/Inspace')
dataset_geocode<-read.csv('dataset_geocoded.csv')

sidewalk_vars<-c('total_num', 'total_sidewalk', 'total_crosswalk')

#create dataset with a column for each variable at each radius level
var.cols<-data.frame(matrix(nrow=nrow(dataset_geocode), ncol=(length(sidewalk_vars)))) #create dataset of columns
colnames(var.cols)<-sidewalk_vars #name the columns

dataset_sidwalk<-cbind(var.cols, dataset_geocode)

```

## 5. Designate ACMT settings

  - Next we set the year and radius for the ACMT
  
```{r designate acmt settings, results='hide'}

#Set the list of variable codes, the list of variable names, the radius, and the year for the data you want pulled
codes_of_acs_variables_to_get<-'B01001_001'
names_of_variables_to_get<-sidewalk_vars
radius_vector <- c(500, 1000, 5000)#set the radius for the area of interest
year <- c(2017) #set the year for the data of interest

```

## 6. Run the ACMT loop to interpolate measures for each lat/long point.

  - Now we can create a loop to pull the Places variables for each location. There is only one year of data to pull from in this dataset.
  

```{r acmt loop, results='hide'}
setwd('/home/rstudio/workspace/')

#run loop to pull variables
for(radius in 1:length(radius_vector)){
radius<-radius_vector[radius]
print(radius)
dataset_radius<-dataset_sidwalk

for(address in 1:nrow(dataset_radius)) {
   tryCatch({if(!is.na(dataset_radius[,1][address])) next #skip the row if the data is already there
  if(!is.na(dataset_radius[,1][address])) next #skip the row if the data is already there
  print(address) #print the number to keep track of progress
  latitude<-dataset_radius$lat[address] #set lat
  longitude<-dataset_radius$long[address] #set long
  
  environmental_measures<-get_acmt_standard_array(long=longitude, lat=latitude, radius_meters = radius, year=year, codes_of_acs_variables_to_get = codes_of_acs_variables_to_get, external_data_name_to_info_list=external_data_name_to_info_list) #pull measures for given lat & long
 
      for(name_of_variable in names_of_variables_to_get){ #for each measures, get the value and put it into the column of the same name
     value_of_variable <- environmental_measures[environmental_measures$names == name_of_variable,]$values  
     dataset_radius[[name_of_variable]][address]<-value_of_variable
  }

 for (name_of_variable in names_of_variables_to_get) {
        dataset_radius[[name_of_variable]][address] <- environmental_measures[environmental_measures$names == name_of_variable, ]$values  
 }},error=function(e){cat("ERROR :", conditionMessage(e), "\n")}) #this will print any error messages
}

if(radius==500) {
dataset_sidewalk2<-dataset_radius
dataset_sidewalk2$radius<-500}

if(radius>500){
dataset_radius$radius<-radius
dataset_sidewalk2<-rbind(dataset_sidewalk2, dataset_radius)
}

}

dataset_sidewalk2<-dataset_sidewalk2%>%
  mutate(year=year)%>% # set year of data pull
  dplyr::select(id, radius, year, everything(), -lat, -long) # reorder variables

```
  
## 7. Calculate sidwalk proportions

```{r calculate proportions}

dataset_sidewalk2<-dataset_sidewalk2%>%
  mutate(prop_sidewalk = total_sidewalk/total_num, 
         prop_crosswalk=total_crosswalk/total_num)

#calculate mean prop_sidewalk and mean prop_crosswalk

mean.sidewalk=mean(dataset_sidewalk2$prop_sidewalk)
mean.crosswalk=mean(dataset_sidewalk2$prop_crosswalk)
sd.sidewalk=sd(dataset_sidewalk2$prop_sidewalk)
sd.crosswalk=sd(dataset_sidewalk2$prop_crosswalk)

#calculate sidewalk and crosswalk z-scores
dataset_sidewalk2 <- dataset_sidewalk2%>%
  mutate(sidewalk_z=(prop_sidewalk-mean.sidewalk)/sd.sidewalk, 
         crosswalk_z=(prop_crosswalk-mean.crosswalk)/sd.crosswalk)

```


## 8. Export dataset

```{rexport data}
setwd('~/workspace/Inspace')
write.csv(dataset_sidewalk2, 'dataset_sidewalk.csv')

```

### References

Nguyen, Q. C., Huang, Y., Kumar, A., Duan, H., Keralis, J. M., Dwivedi, P., ... & Tasdizen, T. (2020). Using 164 million google street view images to derive built environment predictors of COVID-19 cases. International journal of environmental research and public health, 17(17), 6359.

