---
title: "Sidewalk View - Data Pull"
author: "Amy Youngbloom"
date: "11/2/2022"
output: html_document
---

```{r setup, include=FALSE}
# knit this file with the working directory of "workspace/" instead of "workspace/examples"
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, eval=FALSE)
knitr::opts_knit$set(root.dir = '~/workspace')
```

## Sidewalk View 

     Quynh Nguyen at University of Maryland has developed a national sidewalk presence dataset by applying machine learning to Google Street View images to produce tract-level estimates of sidewalk presence/panoramic image by tract. Measures included percent of images for a given census tract that has crosswalks, the percent of images for a given census tract with a sidewalk We will use area-weighted interpolation to construct buffer specific measures of sidewalk presence. 
     
### 1. Setup ACMT

```{r setup acmt}
source('setup-acmt.R')
source('Inspace/Inspace_external_data_functions.R')
```

### 2. Designate settings for data pull

```{r designate settings}
years<-c(2017) #data is only available for 2017
radius_vector<-c(500, 1000, 5000)
variable_list<-names(sidewalk_variable_name_to_interpolate_by_sum_boolean_mapping)
names_of_variables_to_get<-variable_list
external_data_name_to_info_list <- list(sidewalk=external_data_presets_sidewalk)
```
### 3. Import dataset and process

```{r}
dataset_geocoded<-read.csv('~/workspace/Inspace/dataset_geocoded.csv')%>%dplyr::select(id, lat, long)
dataset_sidewalk<-add_rows_columns(dataset_geocoded, variable_list=variable_list, years=years, radius_vector=radius_vector)
```

### 4. Run loop to pull measures

```{r acmt loop, results='hide'}
#run loop to pull variables
for(i in 1:length(years)){
year<-years[i]
print(year)
for(radius in 1:length(radius_vector)){
radius<-radius_vector[radius]
print(radius)

for(address in 1:nrow(dataset_sidewalk[dataset_sidewalk$radius==radius & dataset_sidewalk$year==year,])) {
   tryCatch({if(!is.na(dataset_sidewalk[dataset_sidewalk$radius==radius & dataset_sidewalk$year==year,][,1][address])) next #skip the row if the data is already there
print(address) #print the number to keep track of progress
  latitude<-dataset_sidewalk[dataset_sidewalk$radius==radius & dataset_sidewalk$year==year,]$lat[[address]] #set lat
  longitude<-dataset_sidewalk[dataset_sidewalk$radius==radius & dataset_sidewalk$year==year,]$long[[address]] #set long
  
##pull measures for each lat/long 
  environmental_measures<-get_acmt_standard_array(long=longitude, lat=latitude, radius_meters = radius, year=year, codes_of_acs_variables_to_get = NULL, external_data_name_to_info_list=external_data_name_to_info_list, fill_missing_GEOID_with_zero = TRUE) #pull measures for given lat & long
 
for (name_of_variable in names_of_variables_to_get) {
        dataset_sidewalk[dataset_sidewalk$radius==radius & dataset_sidewalk$year==year,][[name_of_variable]][address] <- environmental_measures[environmental_measures$names == name_of_variable, ]$values 
        
 }},error=function(e){cat("ERROR :", conditionMessage(e), "\n")}) #this will print any error messages
}
}
}

#### Notes: this loop is set up to fill in your dataset as it runs. If the loop is interrupted, you can re-run this chunk only, and it will skip over data that is has already pulled. ####
```

### 5. Calculate proportions & Z-score

```{r calculate proportions}
dataset_sidewalk<-dataset_sidewalk%>%
  mutate(prop_sidewalk = total_sidewalk/total_num, 
         prop_crosswalk=total_crosswalk/total_num)

#calculate mean prop_sidewalk and mean prop_crosswalk

mean.sidewalk=mean(dataset_sidewalk$prop_sidewalk)
mean.crosswalk=mean(dataset_sidewalk$prop_crosswalk)
sd.sidewalk=sd(dataset_sidewalk$prop_sidewalk)
sd.crosswalk=sd(dataset_sidewalk$prop_crosswalk)

#calculate sidewalk and crosswalk z-scores
dataset_sidewalk<- dataset_sidewalk%>%
  mutate(sidewalk_z=(prop_sidewalk-mean.sidewalk)/sd.sidewalk, 
         crosswalk_z=(prop_crosswalk-mean.crosswalk)/sd.crosswalk)

```


### 5. Export dataset

```{r export data}
dataset_sidewalk<-dataset_sidewalk %>% dplyr::select(-lat, -long)

write.csv(dataset_sidewalk, '~/workspace/Inspace/dataset_sidewalk.csv')

```


### References

Nguyen, Q. C., Huang, Y., Kumar, A., Duan, H., Keralis, J. M., Dwivedi, P., ... & Tasdizen, T. (2020). Using 164 million google street view images to derive built environment predictors of COVID-19 cases. International journal of environmental research and public health, 17(17), 6359.

