---
title: "External Data - Walkability"
author: "Amy Youngbloom"
date: "4/8/2022"
output: html_document
---

```{r setup, include=FALSE}
# knit this file with the working directory of "workspace/" instead of "workspace/examples"
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, eval=FALSE)
knitr::opts_knit$set(root.dir = '~/workspace')
```

## EPA Walkability Data

https://www.epa.gov/smartgrowth/smart-location-mapping#walkability

The National Walkability Index is a nationwide geographic data resource that ranks block groups according to their relative walkability. The national dataset includes walkability scores for all block groups as well as the underlying attributes that are used to rank the block groups. The National Walkability Index User Guide and Methodology describes how to use the index and the methodology used to derive the index and ranked scores for its inputs.

### 1. Setup ACMT

```{r setup acmt}
source('setup-acmt.R')
source('Inspace/inspace_external_data_functions.R')
```

### 2. Designate settings for ACMT - NO UPDATES NEEDED

Next we need to designate settings for the ACMT using the walkability presets

```{r, results='hide'}

radius_vector=c(500, 1000, 5000)
years=c(2018)#only 2018 data is available for walkability

external_data_name_to_info_list <- list(
  walkability=external_data_presets_walkability
)

variable_list <-names(walkability_variable_name_to_interpolate_by_sum_boolean_mapping) 
names_of_variables_to_get<-variable_list

```

### 3. Process geocoded dataset
Create additional columns in the dataset for each variable to be pulled and rows for each radii/year

```{r designate acmt settings, results='hide'}

dataset_geocode<-read.csv('Inspace/dataset_geocoded.csv')%>%dplyr::select(id, lat, long)
dataset_walk<-add_rows_columns(dataset=dataset_geocode, variable_list = variable_list,years=years, radius_vector=radius_vector)
```


### 5. Run the ACMT loop to interpolate data for each lat/long

  - Now we can create a loop to pull the Walkability variables for each location, interpolating for each radius

```{r acmt loop, results='hide'}
setwd('~/workspace/')

#run loop to pull variables
for(i in 1:length(years)){
year<-years[i]
print(year)
for(radius in 1:length(radius_vector)){
radius<-radius_vector[radius]
print(radius)

for(address in 1:nrow(dataset_walk[dataset_walk$radius==radius & dataset_walk$year==year,])) {
   tryCatch({if(!is.na(dataset_walk[dataset_walk$radius==radius & dataset_walk$year==year,][,1][address])) next #skip the row if the data is already there
print(address) #print the number to keep track of progress
  latitude<-dataset_walk[dataset_walk$radius==radius & dataset_walk$year==year,]$lat[address] #set lat
  longitude<-dataset_walk[dataset_walk$radius==radius & dataset_walk$year==year,]$long[[address]] #set long
  
##pull measures for each lat/long 
setwd('~/workspace')  
  environmental_measures<-get_acmt_standard_array(long=longitude, lat=latitude, radius_meters = radius, year=year, codes_of_acs_variables_to_get = NULL, external_data_name_to_info_list=external_data_name_to_info_list, fill_missing_GEOID_with_zero = TRUE) #pull measures for given lat & long
 
for (name_of_variable in names_of_variables_to_get) {
        dataset_walk[dataset_walk$radius==radius & dataset_walk$year==year,][[name_of_variable]][address] <- environmental_measures[environmental_measures$names == name_of_variable, ]$values 
        
 }},error=function(e){cat("ERROR :", conditionMessage(e), "\n")}) #this will print any error messages
}
}
}

#### Note: this loop is set up to fill in your dataset as it runs, so if the loop is interrupted, you can re-run this chunk only, and it will skip over data that is has already pulled. ####
```

### 6. Export dataset

Finally, once the measures for each lat/long and each radius has been pulled, we can export and save the dataset

```{r export dataset}
setwd('~/workspace/Inspace')

dataset_walk<-dataset_walk %>% dplyr::select(everything(), -lat, -long)
write.csv(dataset_walk, 'dataset_walkability.csv')
```
