---
title: "External Data - CrimeRisk"
author: "Amy Youngbloom"
date: "11/2/2022"
output: html_document
---

```{r setup, include=FALSE}
# knit this file with the working directory of "workspace/" instead of "workspace/examples"
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, eval=FALSE)
knitr::opts_knit$set(root.dir = '~/workspace')
```

## AGS CrimeRisk

Applied Geographic Systems publishes a block-group level dataset of ‘crime risk’.  This pulls from the FBI’s uniform crime reports and 16000 local law enforcement jurisdictions.  We will use area-weighted interpolation of block group-level estimates to construct buffer-specific estimates of the following measures of reported crime:

## 1. Set up ACMT

```{r setup acmt}
source('setup-acmt.R')
source('Inspace/inspace_external_data_functions.R')
```

### 2. Designate settings for data pull

```{r designate settings}
years<-c(2017) #set year of data pulls (may only be one available)
radius_vector<-c(500, 1000, 5000) #set radii
variable_list<-names(crimerisk_variable_name_to_interpolate_by_sum_boolean_mapping) ## set variables -- should be designated in the external_data_presets
names_of_variables_to_get<-variable_list
external_data_name_to_info_list <- list(crimerisk=external_data_presets_crimerisk) ## update with presets for external dataset (be sure presets are designated in the external_data-presets.R)
```

## 3. Import dataset and process

Next pull in your geocoded dataset and set it up for pulling the Crime Risk data

```{r import geocoded data}
dataset_geocoded<-read.csv('Inspace/dataset_geocoded.csv') %>% dplyr::select(id, lat, long)
dataset_crimerisk<-add_rows_columns(dataset_geocoded, variable_list=variable_list, years=years, radius_vector=radius_vector)

```

## 4. Run loop to pull crime data

```{r acmt loop, results='hide'}

#run loop to pull variables
for(year in 1:length(years)){
year<-years[year]
print(year)
for(radius in 1:length(radius_vector)){
radius<-radius_vector[radius]
print(radius)

for(address in 1:nrow(dataset_crimerisk[dataset_crimerisk$radius==radius & dataset_crimerisk$year==year,])) {
   tryCatch({if(!is.na(dataset_crimerisk[dataset_crimerisk$radius==radius & dataset_crimerisk$year==year,][,1][address])) next #skip the row if the data is already there
print(address) #print the number to keep track of progress
  latitude<-dataset_crimerisk[dataset_crimerisk$radius==radius & dataset_crimerisk$year==year,]$lat[address] #set lat
  longitude<-dataset_crimerisk[dataset_crimerisk$radius==radius & dataset_crimerisk$year==year,]$long[[address]] #set long
  
##pull measures for each lat/long  
  environmental_measures<-get_acmt_standard_array(long=longitude, lat=latitude, radius_meters = radius, year=year, codes_of_acs_variables_to_get = NULL, external_data_name_to_info_list=external_data_name_to_info_list, fill_missing_GEOID_with_zero = TRUE) #pull measures for given lat & long
 

 for (name_of_variable in names_of_variables_to_get) {
        dataset_crimerisk[dataset_crimerisk$radius==radius & dataset_crimerisk$year==year,][[name_of_variable]][address] <- environmental_measures[environmental_measures$names == name_of_variable, ]$values 
        
 }},error=function(e){cat("ERROR :", conditionMessage(e), "\n")}) #this will print any error messages
}
}
}

#### Notes: this loop is set up to fill in your dataset as it runs. If it is interrupted, no problem. You can re-run this chunk only, and it will skip over data that is has already pulled. ####

```
  
## 6. Export final Crime Risk Dataset

```{r export crime data}
dataset_crimerisk<-dataset_crimerisk %>% dplyr::select(-lat, -long)
write.csv(dataset_crime, 'Inspace/dataset_crimerisk.csv')

```
 
