---
title: "mRFEI for Inspace"
author: "Amy Youngbloom"
date: "4/15/2022"
output: html_document
---

```{r setup, include=FALSE}
# knit this file with the working directory of "workspace/" instead of "workspace/examples"
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, eval=FALSE)
knitr::opts_knit$set(root.dir = '~/workspace')
library(tidyverse)
```

## The Modified Food Retail Environment Index

  - The CDC’s modified retail food environment index provides census tract level estimates the ratio of healthy food outlets to total food outlets.  We will use area-weighted interpolation of tract-level estimates to construct buffer-specific estimates of the following measures. 

### 1. Setup ACMT

```{r setup acmt}
source('setup-acmt.R')
source('Inspace/inspace_external_data_functions.R')
```

### 2. Designate settings for data pull

```{r designate settings}
years<-c(2011) #set year of data pulls (may only be one available)
radius_vector<-c(500, 1000, 5000) #set radii
variable_list<-names(mrfei_variable_name_to_interpolate_by_sum_boolean_mapping) ## set variables -- should be designated in the external_data_presets
names_of_variables_to_get<-variable_list
external_data_name_to_info_list <- list(mrfei=external_data_presets_mrfei) ## update with presets for external dataset (be sure presets are designated in the external_data-presets.R)
```

### 3. Import dataset and process

```{r import dataset}
dataset_geocoded<-read.csv('~/workspace/Inspace/dataset_geocoded.csv')%>%dplyr::select(id, lat, long)
dataset_mrfei<-add_rows_columns(dataset_geocoded, variable_list=variable_list, years=years, radius_vector=radius_vector)
```

### 4. Run loop to pull measures

  - Now we can create a loop to pull the mrfei variables for each location
  
```{r acmt loop, results='hide'}

#run loop to pull variables
for(i in 1:length(years)){
year<-years[i]
print(year)
for(radius in 1:length(radius_vector)){
radius<-radius_vector[radius]
print(radius)

for(address in 1:nrow(dataset_mrfei[dataset_mrfei$radius==radius & dataset_mrfei$year==year,])) {
   tryCatch({if(!is.na(dataset_mrfei[dataset_mrfei$radius==radius & dataset_mrfei$year==year,][,1][address])) next #skip the row if the data is already there
print(address) #print the number to keep track of progress
  latitude<-dataset_mrfei[dataset_mrfei$radius==radius & dataset_mrfei$year==year,]$lat[address] #set lat
  longitude<-dataset_mrfei[dataset_mrfei$radius==radius & dataset_mrfei$year==year,]$long[[address]] #set long
  
##pull measures for each lat/long 
setwd('~/workspace')  
  environmental_measures<-get_acmt_standard_array(long=longitude, lat=latitude, radius_meters = radius, year=year, codes_of_acs_variables_to_get = NULL, external_data_name_to_info_list=external_data_name_to_info_list, fill_missing_GEOID_with_zero = TRUE)
 for (name_of_variable in names_of_variables_to_get) {
        dataset_mrfei[dataset_mrfei$radius==radius & dataset_mrfei$year==year,][[name_of_variable]][address] <- environmental_measures[environmental_measures$names == name_of_variable, ]$values 
        
 }},error=function(e){cat("ERROR :", conditionMessage(e), "\n")}) #this will print any error messages
}
}
}

#### Notes: this loop is set up to fill in your dataset as it runs, so if the loop is interrupted, you can re-run this chunk only, and it will skip over data that is has already pulled. ####
```

### 5. Export dataset

```{r export data}
dataset_mrfei<-dataset_mrfei %>% dplyr::select(everything(), -lat, -long)
write.csv(dataset_mrfei, 'Inspace/dataset_mrfei.csv')

```

